Debugging by \textit{lastChange}

### 2010-10-21 trunk/papers/pldi2011/sigplanconf-template.tex

\textit{lastChange} can be called on two different types of values: An object property or a variable value. In JavaScript, every available variable in a frame comes from one of the scopes in the frame's scope chain. There are five different scope types: global, local, with, catch, closure. We explain each case separately.

@@

\subsection{\textit{lastChange} on variable} 
In JavaScript, every frame has a scope chain and every available variable in the frame comes from one of the scopes in the frame's scope chain. Once developer asks for the last change of variable \texttt{foo} at the reproduction point, debugger first determines the variable's scope as follows: it iterates over the scopes in the scope chain and the first scope which has a variable with the same name is the variable's scope. There are five different scope types: local, global, \texttt{with}, closure, \texttt{catch}. We explain these cases in two groups.

@@

A variable in these types of scope can only be changed in the function the variable is defined or the nested fuctions. For example in Figure ~\ref{fig:js-closure}, variable \texttt{var1} in line 11 can only be changed in lines 10 to 23. Therefore, after locating the function in which the variable is defined, it is enough to parse the code inside the function block and set a hook on all lines where the variable is assigned a new value.% TODO lines or instruction???
We also set a hook at the first line of the function which is corresponding to the line variable is defined. In Figure ~\ref{fig:js-closure}, if the execution is paused at line 17 and the last change of \texttt{var1} is queried, two hooks on lines 16 and 17 will be set, but if the execution is paused at line 20 , four hooks on lines 11, 12, 14, 20 will be set. These two cases are different due the fact that \texttt{var1} in \texttt{firstChild} is a local variable but in \texttt{secondChild} is a closure variable.

Everytime one of the hooks hits a new change event is inserted to the list of change events. During execution the same variable might be changed several times but in different scopes. Figure ~\ref{fig:recursive} demonstrates one case that a scope identifier is necessary for locating the right answer. In this case the same variable \texttt{x} is changed several times but in different scopes due to the recursive calls. Without the scope id calling \textit{lastChange(x)} at point C will return point B while the right answer is point A.

@@

@@

To address this issue, debugger does a dependency analysis between all \textit{lastChange} queries and computes the additional data required to be collected at each point. For example in the mentioned case, \texttt{baz} id- if it is available- should be collected at all points \texttt{x} changes. Having this additional data, locating point C, using \texttt{baz} id at point B is straightforward (Figure ~\ref{fig:lastchange-lastchange}).

@@

\FOR{$q$ in $queries$} 
 \FOR {$p$ was defined at the $q$ result}
   \IF {$p$ is a lastChange on object property} 
     \STATE property id should be stored at $q$ change events. 
   \ELSIF {$p$ is a lastChange on variable} 
     \STATE variable scope id should be stored at $q$ change events.
	 \ENDIF 
 \ENDFOR 
\ENDFOR

>>>>

\textit{lastChange} can be called on two different types of values: An object property or a variable value. In JavaScript, every available variable in a frame comes from one of the scopes in the frame's scope chain. We explain each case separately.

@@

\subsection{\textit{lastChange} on variable} 
In JavaScript, every frame has a scope chain and every available variable in the frame comes from one of the scopes in the frame's scope chain. Once developer asks for the last change of variable \texttt{foo} at the reproduction point, debugger first determines the variable's scope as follows: it iterates over the scopes in the scope chain and the first scope which has a variable with the same name is the variable's scope. There are five different scope types: local, global, \texttt{with}, closure and \texttt{catch}. We explain these cases in two groups.

@@

A variable in these types of scope can only be changed in the function the variable is defined or the nested fuctions. For example in Figure ~\ref{fig:js-closure}, variable \texttt{var1} in line 11 can only be changed in lines 10 to 23. Therefore, after locating the function in which the variable is defined, it is enough to parse the code inside the function block and set a hook on all lines where the variable is assigned a new value. % TODO lines or instruction??? 
We also set a hook at the first line of the function which is corresponding to the line variable is defined. In Figure ~\ref{fig:js-closure}, if the execution is paused at line 17 and the last change of \texttt{var1} is queried, two hooks on lines 16 and 17 will be set, but if the execution is paused at line 20 , four hooks on lines 11, 12, 14, 20 will be set. These two cases are different due the fact that \texttt{var1} in \texttt{firstChild} is a local variable but in \texttt{secondChild} is a closure variable.

@@

Everytime one of the hooks hits a new change event is inserted to the list of change events. During execution the same variable might be changed several times but in different scopes. Figure ~\ref{fig:recursive} demonstrates a case that the same variable \texttt{x} is changed several times but in different scopes due to the recursive calls. Without the scope id calling \textit{lastChange(x)} at point C will return point B while the right answer is point A. To address this issue, we employ \texttt{scopeId()} operation-explained in the next section-which returns a unique id for a scopes during on execution. The scope id is kept at every variable change event. Once the execution reaches the reproduction point, using variable's scope id, the last change is recognized.

@@

To address this issue, we use Algorithm \ref{dependency-analysis}, to create the list of additional data should be collected at a change event. This algorithm does does a dependency analysis between all \textit{lastChange} queries and computes the list additional data required to be collected at each change event. For example in the mentioned case, \texttt{baz} id-if it is available- should be collected at all points \texttt{x} changes. Having this additional data, locating point C, using \texttt{baz} id at point B is straightforward (Figure ~\ref{fig:lastchange-lastchange}).

@@

\FOR{$q$ in $queries$} 
 \FOR {$p$ is defined at the $q$ result}
   \IF {$p$ is a lastChange on object property} 
     \STATE the property owner id should be stored at $q$ change events. 
   \ELSIF {$p$ is a lastChange on variable} 
     \STATE the variable scope id should be stored at $q$ change events.
	 \ENDIF 
 \ENDFOR 
\ENDFOR
